### 2020-03-02 test.Rnw

% Rscript -e "library(knitr); knit('test.Rnw')"
% pdflatex test
% biber test
% pdflatex test
% pdflatex test

\documentclass[12pt, a4paper, titlepage]{article}

<<LoadingPackages, include=FALSE>>=
library(knitr)
library(xtable)
library(dplyr)
library(stargazer)
library(cowplot)
library(texreg)
@

\begin{document}

<<Partition, echo=FALSE, results='asis', message=TRUE, warning=TRUE>>=

# first, read in all Rds objects that you need

partition_meals = readRDS("./ANALYSIS/Tables/PartitionMeals.4.Rds")
partition_trips = readRDS("./ANALYSIS/Tables/PartitionTrips.4.Rds")

# colnames(partition_meals) <- c("Variable, Meals", "Mapping, Meals", "Information, Meals")
# colnames(partition_trips) <- c("Variable, Trips", "Mapping, Trips", "Information, Trips")

xtable(partition_meals,
                               # digits = c(0, 0, 1, 1, 0, 0),
          caption = "Partition of Outcomes, Meals", 
          label = 'partitionmeals',
          align = "lccc"
          )
#print(partition_meals, include.rownames = FALSE)

# xtable(partition_trips,
#                                #digits = c(0, 0, 1, 1, 0, 0),
#           caption = "Partition of Outcomes, Trips",
#           label = 'partitiontrips',
#           align = "lccc"
#           )
# print(partition_meals, partition_trips, include.rownames = FALSE)
@

\end{document}